
\chapter*{Preface}
\chaptermark{Preface}
\addcontentsline{toc}{chapter}{Preface}

\vspace{-5mm}
\lettrine[loversize=-0.2,lines=1]{T}{}his dissertation, entitled ``How the Idea of Terrorism Is Changing Us,'' aims to advance existing knowledge on how citizens respond to terrorism, and is based on the following collection of articles:

\begin{itemize}
    \item[A.] Godefroidt, A., \& Langer, A. (2018). How Fear Drives Us Apart: Explaining the Relationship Between Terrorism and Social Trust. \textit{Terrorism and Political Violence}.
    \item[B.] Godefroidt, A., \& Langer, A. (n.d.). Remorse, Repent, and Reconcile: A Conjoint Experiment on Reintegration Attitudes in Nigeria. Revise and Resubmit at \textit{Journal of Peace Research}.
    \item[C.] Demarest, L., Godefroidt, A., \& Langer, A. (2020). Understanding News Coverage of Religious-based Violence: Empirical and Theoretical Insights from Media Representations of Boko Haram in Nigeria. \textit{Journal of Communication}.
    \item[D.] Godefroidt, A. (n.d.). Does Terrorism Really Affect Citizens' Attitudes? A Meta-Analysis. \textit{Under Review.}
\end{itemize}

\noindent These self-contained articles are preceded by an introduction that, first, motivates the questions and overarching theoretical framework guiding this dissertation and, second, summarizes the data and designs used. Then, the four empirical articles are presented. After each article, I quickly take stock of the lessons learned in and questions raised by that article. In the concluding chapter, I take a step back from the data to put all findings into a broader theory- and policy-relevant perspective. In short, this dissertation is an interdisciplinary story---straddling the fields of  social and political psychology, political science, and media studies---about the formation of social and political attitudes in times of terror. Yet, it is also the story of my doctoral journey. As I arranged the articles chronologically, the reader will discover how my way of thinking about this topic has evolved over the course of this dissertation. Let's begin our journey!

\clearpage